Waving and swaying

### 2014-08-27 Paper/grass.tex
Explained using Two Modes instead of KH

%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
\newline
Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Ghisal02}.
Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domain is applicable only to marine canopies. 
\newline   
Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
\newline
While the shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations, several aspect of existing theory remain unsatisfactory. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent theory. Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow. Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of phenomenon.

In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations. Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation. We call this new instability the unvegetated layer instability for brevity. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory with field and experimental observations.

>>>>

%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
\newline
Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically\cite{Vivoni98}. Another major
difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies\cite{Ghisal02}.
Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domain is applicable only to marine canopies. 
\newline   
Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
\newline
While the shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations, several aspect of existing theory remain unsatisfactory. First, linear stability analysis\cite{Raupach96} does not take into account influence of drag due to grass hence is not a self consistent theory. Second, classical free shear flow is known to be unstable for all Reynolds number\cite{drazin}, whereas observations\cite{Grizzle96} indicate the existence of a threshold flow speed below which waving is not observed. Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of the phenomenon.

In this paper we present a mathematical model exhibiting a rich behavior, which is not only consistent within itself, but also with lab experiments and field observations of the threshold. The synchronous waving may be understood to be arising from the linear instability of a uniform flow over the grass bed. In the limit of a dense grass bed, two distinct unstable modes can be asymptotically distinguished, with one of them sharing many characteristics with the Kelvin-Helmholtz instability. In the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.

% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation. We call this new instability the unvegetated layer instability for brevity. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory with field and experimental observations.
